Molecular dynamics simulation of synchronization in driven particles

### 2021-01-21 manuscript/manuscript_ajp_2020.tex
updated intro into a single section, started generalizing the conclusion

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
In Section~\ref{sec:colloids} we describe colloid particles,
an relevant
experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
In Section~\ref{sec:conclusion}
we describe how our work with colloids applies more broadly
to many physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

\section{Colloidal Models}
\label{sec:colloids}

@@

Light is a
tool for manipulating the colloidal environment.
A single laser beam can be used to create an optical trap,
which uses 
photon radiation pressure to
control colloid motions \cite{Ashkin1997}.
A circular optical trap is designed
with a local energy minimum 
at its center.
A colloid in the beam center is subject to a
uniform distribution of photon bombardment,
and motions away from the center create a net force
due to an uneven 
distribution of photon collisions on its surface.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either restores colloids toward the local minima
or ejects it from the trap.
A single optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.

In experiments  
 complex environments can be
 engineered with light by %, typically made 
 using diffraction gratings to create a patterned
 potential energy landscape 
 with one or more laser beams \cite{}.
Experiments of colloid particles confined
in optical traps arrays or patterns 
subject to external driving forces 
have
been used to examine the microscopic dynamics
of individual particles during mode-locking \cite{Juniper2015}.
This produced step-by-step diagrams
of particle dynamics that cannot be 
imaged in electronic systems such as Josephson junctions.
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Thus the study of
mode-locking is a useful problem
of complex quantum mechanical systems.

%external forces
%The
%dynamics of particles subject to 
%an applied external force apply to many physical systems.
%For instance the flow of charges in a conductor,
%or cells responding to a chemical gradient.
%The external force 
%increases the diversity of dynamical behaviors,
%and can cause particles to flow in
%a variety of non-linear complex patterns
%that may be 
%synchronized.
%Disordered chaotic dynamics are also possible,
%where irregular, unpredictable time evolution of
%nonlinear systems occurs in mechanical oscillators \cite{chaos}.
%-------------------------------------------------------------

Numerical modeling of colloids can provide mechanistic insight
that can be difficult to achieve in experimental conditions
where Brownian motion and other sources of noise dominate.

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
%Fundamental models
%of synchronization 
%can %include many coupled oscillators
%or
%focus on 
A single particle driven across a periodic potential landscape 
%Single particle studies are useful
is useful 
to study synchronization
in the absence of collective effects of particle interactions.
%due to interactions between particles.
%A single particle particle oscillator
%within a potential well is 
%much like a skateboarder in a half-pipe or
%a child on a swing
%moving back and forth. % through equilibrium.  % \cite{}.
The relationship between the driving force and
landscape can change the pattern of motion.  
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning \cite{Reichhardt2017}.  
%and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.
Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE

Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.
%Oscillators ordered into chains or lattices.
%has many important
%technological applications
Dynamical mode-locking
is %often
observed
in quantum electronic
devices such as Josephson junctions \cite{Josephson1962,Josephson1965}.
A single junction contains 
two superconductoring layers which sandwich an insulating layer.
When subject to an external voltage,
Cooper pairs in the superconducting materials
tunnel through the insulating layer.
Phase-locking is observed as 
%phase-locked current loops \cite{} 
stepped regions in current-voltage (I-V) relationship in these devices,
where voltage is the analog of external driving force
and current is that of particle velocity.
%where a constant current results from a range of voltages
%due to the quantized nature of the tunneling probability [NOT QUITE RIGHT...].
%instead of ohmic relationships.
Known as Shapiro steps, % \cite{},
these phase locked currents  
have been observed due to applied ac voltages in 
single Josephson junctions \cite{Shapiro1963, Golubov2004} and
coupled arrays of junctions \cite{Benz1990}.
%either in 
%velocity-force curves and voltage-current curves \cite{Golubov2004}.
Shapiro steps vary in width depending on the strength of the
applied ac forces,
and are observed in a variety of systems
displaying
%are hallmarks of
non-Ohmic behavior in voltage-current curves,
including
ac and dc driven
charge and spin density waves
and superconducting vortices in landscapes 
engineered with periodic patterns of pinning sites.
Our results can be relevant 
to synchronization effects
in a broad range of experiments systems
including optically confined colloids,
superconductors with periodic pinning arrays, 
and the charge and spin of atomic systems.

>>>>

%\section{Colloidal Models}
%\label{sec:colloids}

@@

Light is a
tool for manipulating the colloidal environment.
A single laser beam can be used to create an optical trap,
which uses 
photon radiation pressure to
control colloid motions \cite{Ashkin1997}.
A circular optical trap is designed
with a local energy minimum 
at its center.
A colloid in the beam center is subject to a
uniform distribution of photon bombardment,
and motions away from the center create a net force
due to an uneven 
distribution of photon collisions on its surface.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either restores colloids toward the local minima
or ejects it from the trap.
An optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.

%In experiments  
 Complex environments can be
 engineered with light by %, typically made 
 using diffraction gratings to create a patterned
 potential energy landscape 
 with one or more laser beams \cite{Grier2003}.
 How a particle moves
 in an optical landscape can
 change
 when 
 subject to external driving forces.
 Individual particles can
 perform 
 synchronized % with the environment
 motions
 that lock together particle location and velocity.
 Experimental studies of
 the microscopic dynamics 
 colloids performing phase-locked motions
 produce step-by-step diagrams
 of particle dynamics useful for
 understanding synchronization at a single particle level \cite{Juniper2015}.

%Numerical modeling of colloids can provide mechanistic insight
%that can be difficult to achieve in experimental conditions
%where Brownian motion and other sources of noise dominate.
%external forces
%The
%dynamics of particles subject to 
%an applied external force apply to many physical systems.
%For instance the flow of charges in a conductor,
%or cells responding to a chemical gradient.
%The external force 
%increases the diversity of dynamical behaviors,
%and can cause particles to flow in
%a variety of non-linear complex patterns
%that may be 
%synchronized.
%Disordered chaotic dynamics are also possible,
%where irregular, unpredictable time evolution of
%nonlinear systems occurs in mechanical oscillators \cite{chaos}.
%-------------------------------------------------------------

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
%In Section~\ref{sec:colloids} we describe colloid particles,
%an relevant
%experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
%In Section~\ref{sec:sync}
%we 
%extend the model to include multiple particles,
%and show how confined 
%interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.
In Section~\ref{sec:conclusion}
we describe how our work with colloids applies more broadly
to many physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
A single particle oscillator
in a potential well is 
much like a skateboarder in a half-pipe or
a child on a swing.
If subject to an applied force,
a confined oscillator may 
synchronize its location
to the periodic pattern of the external drive,
moving back and forth in time with
the beat. 
%Fundamental models
%of synchronization 
%can %include many coupled oscillators
%or
%focus on
We have used 
a single particle driven across a periodic potential landscape 
%Single particle studies are useful
to study synchronization dynamics. 
%due to interactions between particles.
The relationship between the driving force and
landscape can change the pattern of motion.  
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning \cite{Reichhardt2017}.  
%and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.
%Oscillators ordered into chains or lattices.
%has many important
%technological applications
Dynamical mode-locking
is %often
observed
in quantum electronic
devices such as Josephson junctions \cite{Josephson1962,Josephson1965}.
A single junction contains 
two superconductoring layers which sandwich an insulating layer.
When subject to an external voltage,
Cooper pairs in the superconducting materials
tunnel through the insulating layer.
Phase-locking is observed as 
%phase-locked current loops \cite{} 
stepped regions in current-voltage (I-V) relationship in these devices,
where voltage is the analog of external driving force
and current is that of particle velocity.
%where a constant current results from a range of voltages
%due to the quantized nature of the tunneling probability [NOT QUITE RIGHT...].
%instead of ohmic relationships.
Known as Shapiro steps, % \cite{},
these phase locked currents  
have been observed due to applied ac voltages in 
single Josephson junctions \cite{Shapiro1963, Golubov2004} and
coupled arrays of junctions \cite{Benz1990}.
%either in 
%velocity-force curves and voltage-current curves \cite{Golubov2004}.
Shapiro steps vary in width depending on the strength of the
applied ac forces,
and are observed in a variety of systems
displaying
%are hallmarks of
non-Ohmic behavior in voltage-current curves,
including
ac and dc driven
charge and spin density waves
and superconducting vortices in landscapes 
engineered with periodic patterns of pinning sites.
Thus the study of
mode-locking is a useful problem
of complex quantum mechanical systems.
Our results can be relevant 
to synchronization effects
in a broad range of experiments systems
including optically confined colloids,
superconductors with periodic pinning arrays, 
and the charge and spin of atomic systems.
